A Toolbox for Manipulating and Assessing Color Palettes for Statistical Graphics

### 2019-03-18 pkg/colorspace/paper/Toolbox/article.Rmd
re-introduce changes from Claire from r466 that were erroneously undone in r469
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/colorspace@492 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

Using the physiologically-based model for simulating color vision deficiency (CVD)
of @color:Machado+Oliveira+Fernandes:2009 different kinds of limitations can be
emulated: deuteranope (green cone cells defective), protanope (red cone cells defective),
and tritanope (blue cone cells defective).

Below we briefly describe our \proglang{R} interface to these emulation techniques
and show them in practice for a heatmap with sequential palette. Another example
with diverging palette is available at
<http://colorspace.R-Forge.R-project.org/articles/color_vision_deficiency.html>.
Finally, CVD emulation is particularly useful for bringing out why the RGB rainbow
palette is almost always a bad choice in scientific displays. See
<http://colorspace.R-Forge.R-project.org/articles/endrainbow.html> for further
illustrations.

@@

The workhorse function to emulate color-vision deficiency is `simulate_cvd()`, which can take any vector
of valid \proglang{R} colors and transform them according to a certain CVD transformation
matrix and transformation equation. The transformation matrices have been established
by @color:Machado+Oliveira+Fernandes:2009 and are provided in objects `protanomaly_cvd`,
`deutanomaly_cvd`, and `tritanomaly_cvd`.
The convenience interfaces `deutan()`, `protan()`, and `tritan()` are the high-level functions for
simulating the corresponding kind of color blindness with a given severity (calling
`simulate_cvd()` internally).

For further guidance on color blindness in relation to statistical graphics
see @color:Lumley:2006 which accompanies the \proglang{R} package \pkg{dichromat} [@color:dichromat]
and is based on earlier emulation techniques
[@color:Vienot+Brettel+Ott:1995; @color:Brettel+Vienot+Mollon:1997; @color:Vienot+Brettel+Mollon:1999].

>>>>

Different kinds of limitations can be emulated using the physiologically-based model
for simulating color vision deficiency (CVD) of @color:Machado+Oliveira+Fernandes:2009:
deuteranomaly (green cone cells defective),
protanomaly (red cone cells defective), and
tritanomaly (blue cone cells defective).
While most other CVD simulations handle only dichromacy, where one of three cones is
non-functional, @color:Machado+Oliveira+Fernandes:2009 provide a unified model of
both dichromacy and anomalous trichromacy, where one cone has shifted spectral
sensitivity. As anomalous trichromacy is the most common form of color vision
deficiency, it is important to emulate along with the rarer, but more severe
dichromacy. Below we briefly describe our \proglang{R} interface to these emulation
techniques and show them in practice for a heatmap with sequential palette.
Another example with diverging palette is available at
<http://colorspace.R-Forge.R-project.org/articles/color_vision_deficiency.html>.
Finally, CVD emulation is particularly useful for bringing out why the RGB
rainbow palette is almost always a bad choice in scientific displays. See
<http://colorspace.R-Forge.R-project.org/articles/endrainbow.html> for further
illustrations.

@@

The workhorse function to emulate color vision deficiencies is `simulate_cvd()`
which can take any vector of valid \proglang{R} colors and transform them according to a
certain CVD transformation matrix and transformation equation. The
transformation matrices have been established by @color:Machado+Oliveira+Fernandes:2009 and are
provided in objects `protanomaly_cvd`, `deutanomaly_cvd`, and `tritanomaly_cvd`.
The convenience interfaces `deutan()`, `protan()`, and `tritan()` are the high-level
functions for simulating the corresponding kind of color blindness with a given
`severity` (calling `simulate_cvd()` internally). A severity of `1` corresponds to
dichromacy, `0` to normal color vision, and intermediate values to varying
severities of anomalous trichromacy. For further guidance on color blindness in
relation to statistical graphics see @color:Lumley:2006 which accompanies the \proglang{R}
package \pkg{dichromat} [@color:dichromat] and is based on earlier emulation techniques
[@color:Vienot+Brettel+Ott:1995; @color:Brettel+Vienot+Mollon:1997; @color:Vienot+Brettel+Mollon:1999].
